### 2020-05-29 notebook/import.Rmd
work on store

# Import

```{r birds}
bench_time(
  neon_download("DP1.10003.001", file_regex = ".*basic.*[.]csv")
)
```

```{r beetles}
bench::bench_time(
  neon_download("DP1.10022.001")
)
```

```{r ticks}
bench::bench_time(
  neon_download("DP1.10093.001", quiet = TRUE)
)
```

```{r mammals}
bench::bench_time(
  neon_download("DP1.10072.001",  quiet = TRUE)
)
```

```{r zooplankton}
bench::bench_time(
  neon_download("DP1.20219.001",  quiet = TRUE)
)
```

```{r macroinvert}
bench::bench_time(neon_download("DP1.20120.001", quiet = TRUE))
```

```{r mozzies}
bench::bench_time(
  neon_download("DP1.10043.001", quiet = TRUE)
)
```

# Explore

```{r}
library(dplyr)

meta <- neon_index()
meta %>% count(product)
```

```{r}
meta %>% count(name)
```

```{r}

bench::bench_time({
f <- meta %>% filter(name == "brd_countdata", type == "basic") %>% pull(path)
df <- neon_read(f) # vroom with error handler
})

```

```{r}

bench::bench_time({
f <- meta %>% filter(name == "mam_perplotnight", type == "basic") %>% pull(path)
df <- neon_read(f) # vroom with error handler
})

```

```{r}
as_tibble(df)
```

# Distribute

```{r}
library(contentid)
library(fs)
library(dplyr)

```

```{r}
meta <- neon_index()
d <- fs::dir_info(neon_dir())
```

how much data?

```{r}
dim(d)[[1]]
d %>% summarise(sum(size))
```

```{r}
ids <- contentid::store(meta$path, dir = "/zpool/content-store")
```