### 2020-06-04 notebook/import.Rmd
support hash validation

# Import

# Checksum comparison

The NEON API returns 32 character crc32 checksum

```{r}
raw_files <- neon_index(table = "brd_countdata-expanded")
x <- raw_files$path[[1]]
digest::digest(x, algo = "crc32", file = TRUE)

url <- paste("https://data.neonscience.org/api/v0/data", 
                raw_files$product[[1]], 
                "BART", 
                raw_files$month[[1]], 
                sep = "/")
resp <- httr::GET(url)
out <- jsonlite::fromJSON(httr::content(resp, "text"))$data$files
out[out$name == basename(x),]

dir.create("tmp")
paste0("tmp/", out$name)
purrr::walk2(out$url, paste0("tmp/", out$name), curl::curl_download)
unzip("tmp/NEON.D01.BART.DP1.10003.001.2015-06.expanded.20191107T154457Z.zip",
      exdir = "tmp/expanded")
```

```{r}
vapply(fs::dir_ls("tmp", glob="*.csv"), digest::digest,  character(1L), algo = "crc32", file = TRUE, USE.NAMES = FALSE)
vapply(fs::dir_ls("tmp/expanded", glob="*.csv"), digest::digest, character(1L), algo = "crc32", file = TRUE, USE.NAMES = FALSE)

out$crc32[grepl("[.]csv",out$name)]

```

# Distribute

```{r}
library(contentid)
library(fs)
library(dplyr)

```

```{r}
meta <- neon_index()
d <- fs::dir_info(neon_dir())
```

how much data?

```{r}
dim(d)[[1]]
d %>% summarise(sum(size))
```

```{r}
ids <- contentid::store(meta$path, dir = "/home/content-store")
```

```{r}
registry <- vroom::vroom("content_id.tsv")

meta <- dplyr::left_join(meta, registry) %>% as_tibble()
vroom::vroom_write(meta, "neonstore.tsv")

```

```{r}
## read from the content-store instead
meta <- vroom::vroom("neonstore.tsv")
metab <- meta %>%  mutate(store = retrieve(id, "/home/content-store"))
```

```{r}
birds <- metab %>% filter(name == "brd_countdata", type == "basic") %>% pull(store) %>% neon_read()

```

```{r}
metab <- metab %>% mutate(mirror = gsub("/home/content-store", "https://data.carlboettiger.info", store))

request <- paste0("https://hash-archive.jetstream.carlboettiger.info/api/enqueue/", metab$mirror)
request2 <- paste0("https://hash-archive.org/api/enqueue/", metab$mirror)
x <- lapply(request2, httr::GET)

#xml2::url_escape("https://data.carlboettiger.info/data/10/c9/10c9e39af827ecdba4255ded08ab7f8e2b6895ce58561236d1e26ff3fa097d20")
```