### 2020-06-08 README.Rmd
:pen: more context about motivations

---
output: 
  github_document:
    df_print: tibble
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
library(neonstore)
Sys.setenv("NEONSTORE_HOME" = tempfile())
```

# neonstore

<!-- badges: start -->
[![R build status](https://github.com/cboettig/neonstore/workflows/R-CMD-check/badge.svg)](https://github.com/cboettig/neonstore/actions)
[![Codecov test coverage](https://codecov.io/gh/cboettig/neonstore/branch/master/graph/badge.svg)](https://codecov.io/gh/cboettig/neonstore?branch=master)
<!-- badges: end -->

`neonstore` provides quick access and persistent storage of NEON data tables. 
`neonstore` emphasizes simplicity and a clean data provenance trail, see 
(Provenance)[/#Provenance].

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cboettig/neonstore")
```
## Quickstart

Discover data products of interest:

```{r}
products <- neon_products()

i <- grepl("Populations", products$themes)
products[i, c("productCode", "productName")]

i <- grepl("bird", products$keywords)
products[i, c("productCode", "productName")]
```

Download all data files in the bird survey data products.

```{r}
library(neonstore)
neon_download("DP1.10003.001")

```

Now, view your store of NEON products:

```{r }
neon_store()
```

These will persist between sessions, so you only need to download once
or to retrieve updates. `neon_store()` can take arguments to filter by product
or pattern in table name, e.g. `neon_store(table = "brd")`.

Once you determine the table of interest, you can
read in all the component tables into a single `data.frame`

```{r}
neon_read("brd_countdata-expanded")
```

Two other functions access additional API endpoints 
that may also be of interest. `neon_sites()` returns a `data.frame`
of site information, including site descriptions and 
the ecological domain that each site falls into:

```{r}
neon_sites()
```

Lastly, `neon_products()` returns a table with a list of all neon products,
which may be useful for data discovery or additional metadata about any
given product:

```{r}
neon_products()
```

## Design Details / comparison to `neonUtilities`

`neonstore` is not meant as a replacement to the `neonUtilities` package
developed by NEON staff.  `neonUtilities` performs a range of product-specific
data querying, parsing, and data manipulation beyond what is provided by NEON's
API or web interface. `neonUtilities` also provides other utilities for working 
with NEON data beyond the scope of the NEON API or the data download/ingest 
process. While this processing is undoubtedly useful, it may make
it difficult to compare results or analyses based on data downloaded and accessed
using `neonUtilities` R package with analyses based on data accessed directly
from the web interface, the API, or another tool (or even a different release
of the `neonUtilities`).

By contrast, `neonstore` aims to do far less.  `neonstore` merely automates the
download of individual NEON data files.  In contrast to `neonUtilities` which 
by default "stacks" these raw files into single tables and discards the raw 
data, `neonstore` preserves only the raw files in the store, stacking the 
individual tables "on demand" using `neon_read()`. `neon_read()` is a thin
wrapper around the `vroom` package, 
[Hester & Wickham, 2020](https://vroom.r-lib.org), which uses the `altrep` 
mechanism in R to provide very fast reads of rectangular text data into R,
and trivially handles the case of a single table being broken across many files.
Some NEON tables are not entirely consistent in their use of columns across the
individual site-month files, so `neon_read()` transparently checks for this,
reading in groups of files sharing all matching columns with `vroom` before
binding the groups together. This makes it easier to always trace an analysis
back to the original input data, makes it easier to update input data files 
without facing the challenge of either downloading & stacking the whole 
data product from scratch again or having to keep track of some previously
downloaded data file.

A few other differences are also worth noting.

- `neonstore` aims to provide persistent storage, writing raw data files to
  the appropriate app directory for your operating system (see `rappdirs`,
  [Ratnakumar et al 2016](https://CRAN.R-project.org/package=rappdirs)). 
  More details about this can be found in [Provenance](/#Provenance), below.
- `neon_download()` provides  clean and concise progress bars for the two key
  processes involved: querying the API to obtain download URLs (which involves no 
  large data transfer but counts against API rate limiting, see below), and the
  actual file downloads.
- `neon_download()` will verify the integrity of file downloads against the MD5
  hashes provided. 
- `neon_download()` will omit downloads of any existing data files in the local 
  store.  
- You can request multiple products at once using vector notation, though API
  rate limiting may interfere with large requests.
- `neon_download()` uses `curl::curl_download()` instead of `downloadr` package
  used in `neonUtilities`, which can be finicky on Windows and older versions of R. 
- `neonstore` has slightly lighter dependencies: only `vroom` and `httr`, and
  packages already used by one of those two (`curl`, `openssl`).

Like `neonUtilities`, You can optionally include site and date filters,
e.g. to request only records more  recent than a certain date.  Doing so will
preserve API quota and improve speed (see API limits, below). `neonUtilities`
is also far more widely tested and has extensive error handling tailored to
individual data products.

## Provenance

Because `neonstore` only stores raw data products as returned from the NEON API,
it can easily determine which files have already been downloaded, and only
download new files without requiring the user to specify specific dates. 
(It must still query the API for all the metadata in the requested date range).
This same modular approach also makes it easy to track _data provenance_, an 
essential element of reproduciblity in comparing results across other analyses
of the NEON data.

We can list precisely which component files are being read in by `neon_read()`
by consulting `neon_index()`:

```{r}
raw_files <- neon_index(table = "brd_countdata-expanded")
raw_files
```

`neon_read()` is a relatively trivial function that simply passes this file 
list to `vroom::vroom()`, a fast, vectorized parser that can easily read in 
a single table that is broken into many separate files.

Imagine instead that we use the common pattern of downloading
these raw files, stacks and possibly cleans the data, saving only this derived
product while discarding the individual files.  Now imagine a second researcher,
at some later date, queries the API over the same reported range of dates and 
sites, uses the same software package to stack the tables, only to discover the
resulting table is somehow different from ours (e.g. by comparing file hashes).
Pinpointing the source of the discrepancy would be challenging and 
labor-intensive.

In contrast, the same detective-work would be easy with the `neonstore` file
list.  We can confirm if the API had returned the same number of 
raw files with the same names; and better, can verify integrity of the contents
by comparing hashes of files now being returned to those recorded
by `neon_index()`.  In this way, we could determine if any additional files
had been included or pinpoint any files that may have changed.

As such, users might want to store the `neon_index()` `data.frame` for the
table(s) they have used as part of their analysis, including the individual
file hashes.  One can also generate a zip of all the data files for 
archival purposes. (Note that NEON is an Open Data provider, see
[LICENCE](https://www.neonscience.org/data/about-data/data-policies).)

```{r message=FALSE, eval=FALSE}
write.csv(raw_files, "index.csv")
zip("brd_countdata.zip", raw_files$path)
```

## Data citation

Always remember to cite your data sources!
`neonstore` knows how to generate the appropriate citation for the data
in your local store (or any specific product).

```{r}
neon_citation()
```

## Note on API limits

[The NEON API now rate-limits requests.](https://data.neonscience.org/data-api/rate-limiting/#api-tokens).
Using a personal token will increase the number of requests you can make.
See that link for directions on registering for a token.
Then pass this token in `.token` argument of `neon_download()`,
or for frequent use, add this token as  an environmental variable, `NEON_DATA`
to your local `.Renviron` file in your user's home directory.

`neon_download()` must first query each the API of eacn NEON site which collects
that product, for each month the product is collected.
(It would be much more efficient on the NEON server if the API could take
queries of the from `/data/<product>/<site>`, and pool the results, rather than
require each month of sampling separately!)

```{r include=FALSE}
unlink("brd_countdata.zip")
unlink("index.csv")
Sys.unsetenv("NEONSTORE_HOME")
```