### 2020-10-06 README.Rmd
update README :pen:

---
output: 
  github_document:
    df_print: tibble
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE
)
library(neonstore)
Sys.setenv("NEONSTORE_HOME" = tempfile())
```

# neonstore

<!-- badges: start -->
[![R build status](https://github.com/cboettig/neonstore/workflows/R-CMD-check/badge.svg)](https://github.com/cboettig/neonstore/actions)
[![Codecov test coverage](https://codecov.io/gh/cboettig/neonstore/branch/master/graph/badge.svg)](https://codecov.io/gh/cboettig/neonstore?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/neonstore)](https://CRAN.R-project.org/package=neonstore)
<!-- badges: end -->

`neonstore` provides quick access and persistent storage of NEON data tables. 
`neonstore` emphasizes simplicity and a clean data provenance trail, see 
Provenance section below.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cboettig/neonstore")
```
## Quickstart

```{r}
library(neonstore)
```

Discover data products of interest:

```{r}
products <- neon_products()

i <- grepl("Populations", products$themes)
products[i, c("productCode", "productName")]

i <- grepl("bird", products$keywords)
products[i, c("productCode", "productName")]
```

Download all data files in the bird survey data products.

```{r}
neon_download("DP1.10003.001")

```

View your store of NEON products:

```{r }
neon_index()
```

These files will persist between sessions, so you only need to download once
or to retrieve updates. `neon_index()` can take arguments to filter by product
or pattern (regular expression) in table name, e.g. `neon_index(table = "brd")`.

Once you determine the table of interest, you can
read in all the component tables into a single `data.frame`

```{r}
neon_read("brd_countdata-expanded")
```

## Database backend

`neonstore` now supports a backend relation database as well. Import data from
the raw downloaded files using `neon_store()`:

```{r}
neon_store(table = "brd_countdata-expanded")
```

Alternately, we could import all data tables associated with a given product:

```{r}
neon_store(product = "DP1.10003.001")
```

Access an imported table using `neon_table()` instead of `neon_read()`:

```{r}
neon_table("brd_countdata-expanded")
```

Access the remote database using `neon_db()`.  This is a `DBIConnection` that
can easily be used with `dplyr` functions like `tbl()` or `filter()`.  
Remember that `dplyr` translates these into SQL queries that run directly
on the database.

```{r}
library(dplyr)

con <- neon_db()
brd <- tbl(con, "brd_countdata-expanded")
brd %>% filter(siteID == "ORNL")

```

## Note on API limits

If `neon_download()` exceeds the API request limit (with or without the token),
`neonstore` will simply pause for the required amount of time to avoid 
rate-limit-based errors.

[The NEON API now rate-limits requests.](https://data.neonscience.org/data-api/rate-limiting/#api-tokens).
Using a personal token will increase the number of requests you can make before
encountering this delay.  See link for directions on registering for a token.
Then pass this token in `.token` argument of `neon_download()`,
or for frequent use, add this token as  an environmental variable, `NEON_DATA`
to your local `.Renviron` file in your user's home directory. 
`neon_download()` must first query each the API of each NEON site which collects
that product, for each month the product is collected.

(It would be much more efficient on the NEON server if the API could take
queries of the from `/data/<product>/<site>`, and pool the results, rather than
require each month of sampling separately!)

```{r include=FALSE}
unlink("my_neondata.zip")
unlink("index.csv")
Sys.unsetenv("NEONSTORE_HOME")
```